Parallel Level-Set Methods on Adaptive Tree-Based Grids

### 2016-02-12 paper/tex/parallel_algorithms.tex
added accuracy section with enright test

Once buffering is done, every process knows exactly how many messages it needs to send and to which processes. This also implicitly defines processes that will later on send a reply message to this process. However, at this point no process knows which processes to expect a message from. We solve this problem using a simple \textit{communication matrix} (see Figure \ref{fig:communication}). Other approaches for solving this problem could include using non-blocking collective or one-sided communication operations introduced in the MPI-3 standard \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication}. Although these algorithms have better theoretical communication complexities, we did not see any difference in the runtime or scalability of algorithm \ref{alg:interpolation} when using them.

@@

>>>>

Once buffering is done, every process knows exactly how many messages it needs to send and to which processes. This also implicitly defines processes that will later on send a reply message to this process. However, at this point no process knows which processes to expect a message from. We solve this problem using a simple \textit{communication matrix} (see Figure \ref{fig:communication}). Our approach is very similar to the ``Personalized Census ($\mathcal{PCX}$)'' algorithm described in \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication}. Another similar approach is the ``Notify'' algorithm introduced in \cite{Isaac;Burstedde;Ghattas:12:Low-cost-parallel-al}. Furthermore, the MPI-3 standard introduces non-blocking collectives and Remote Memory Access (RMA) operations which enable new ways of solving the communication problem. For instance, authors in \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication} descried the ``Non-blocking Consensus ($\mathcal{NBX}$)'' and ``Remote Summation ($\mathcal{RSX}$)'' algorithms which make use of such operations and have better theoretical communication complexities. With the exception of $\mathcal{RSX}$ algorithm, which was not tested in this study, all remaining algorithms produced similar timing and scaling. Thus we have decided to describe our algorithm based on the idea of the \textit{communication matrix}.

@@

\subsection{Accuracy} \label{sec:accuracy}

The numerical methods detailed in the previous sections are widely used in the level-set community and their accuracy is studied for example in \cite{Min;Gibou:07:A-second-order-accur}. However, in order to validate our implementation, we present a brief convergence analysis.

The advection of an irregular boundary using the semi-Lagrangian method is the perfect candidate to demonstrate the accuracy of our implementation as it makes use of the interpolation routine as well as the reinitialization procedure. We select the benchmark problem proposed in \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}. Consider a sphere centered at $(0.35,0.35,0.35)$ and with radius $0.15$ in a domain $[0,1]^3$ and deformed under the divergence free velocity field
\be \label{eq:velo_enright}
\begin{cases}
u(x,y,z) = 2 \sin^2(\pi x)\sin(2 \pi y)\sin(2 \pi z), \\
v(x,y,z) = - \sin^2(\pi y)\sin(2 \pi x)\sin(2 \pi z), \\
w(x,y,z) = - \sin^2(\pi z)\sin(2 \pi x)\sin(2 \pi y).
\end{cases}
\ee
forward in time until $t_{1/2}=1$ and then backward to its original state at $t_f=2$. With set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest. We monitor the volume loss and the error in the interface location at the final time, when the original shape should be recovered, as the finest resolution of the forest increases. The results are reported in table \ref{tab:accuracy} and figure \ref{fig:accuracy} shows a visualization of the sphere at the initial time, at $t_{1/2}$ when the deformation is maximal, and at the final time. The results are consistent with those reported previously in \cite{Min;Gibou:07:A-second-order-accur} and indicate second order accuracy for the mass loss.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Finest resolution & $L^\infty$ error on $\phi$ & Rate & Volume loss (\%) & Rate \\
\hline
$128^3$  & 1.53E-01 & -    & 1.85E-01 & -    \\
$256^3$  & 1.18E-01 & 0.38 & 3.73E-02 & 2.31 \\
$512^3$  & 9.82E-03 & 3.58 & 6.87E-03 & 2.44 \\
$1024^3$ & 2.56E-03 & 1.94 & 1.92E-03 & 1.84 \\
$2048^3$ & 1.60E-03 & 0.67 & 5.17E-04 & 1.89 \\
\hline
\end{tabular}
\caption{Study of the convergence of the level-set algorithm using Enright's test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}. The $L^\infty$ error on $\phi$ is computed only close to the interface as this is the relevant observable for an advection procedure.} \label{tab:accuracy}
\end{center}
\end{table}

\begin{figure}[htbp]
\begin{center}
\begin{tabular}{ccccc}
\includegraphics[width=.28\textwidth]{figures/enright_0.png}
&
\hspace{.1cm}
&
\includegraphics[width=.28\textwidth]{figures/enright_1.png}
&
\hspace{.1cm}
&
\includegraphics[width=.28\textwidth]{figures/enright_2.png}
\end{tabular}
\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$ and colored with the processes ranks. The forest's finest resolution is equivalent to a uniform grid with $2048^3$ cells and contains between 17 (at initial and final times) and 63 (at $t_{1/2}$) million nodes. These results were obtained on the Comet supercomputer with 960 processes (40 compute nodes).} \label{fig:accuracy}
\end{center}
\end{figure}

### 2016-02-12 paper/tex/scaling_results.tex
added accuracy section with enright test

\subsection{Semi-Lagrangian}
To test the scalability of the semi-Lagrangian scheme of Algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e. we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by:
\bean
u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
\eean
To understand the effect of the CFL number on the scalability of the algorithm we perform one step of the semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. We also perform the test for two different initial girds, a small grid with maximum level $l_\text{max} = 10$ and a large grid with maximum level $l_\text{max} = 12$. In both cases, the minimum level is $l_\text{min}=0$. After one advection step, these grids have approximately 15M and 255M nodes, respectively. 
\begin{table}[htbp]
	\begin{minipage}{.48\linewidth}
		\begin{center}
		\begin{tabular}{|c|cccccc|}
			\hline
				\diagbox{CFL}{\#p} & 16 & 32 & 64 & 128 & 256 & 512 \\
				\hline
				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
			\hline
		\end{tabular}
		\caption*{(a) $l_\text{max} = 10$}
		\end{center}
	\end{minipage}%
	\begin{minipage}{.48\linewidth}
		\begin{center}
		\begin{tabular}{|c|cccccc|}
			\hline
				\diagbox{CFL}{\#p} & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
			\hline
				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
			\hline
		\end{tabular}
		\caption*{(b) $l_\text{max} = 12$}
		\end{center}
	\end{minipage}	
	\caption{Number of sub-iterations required for the grid construction in Algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively. Note how the sub-iteration count increases with the CFL number but is almost independent of the number of processes. The slight dependence between the number of sub-iterations and the number of processes is most likely due to the dependence of round-off errors on the number of processes. Nonetheless, close examination of the Octrees generated (data not shown) reveals that they are identical and independent of the number of processes used to perform the test.}
	\label{tab:semilagrangian}
\end{table}

>>>>

\subsection{Semi-Lagrangian}
To test the scalability of the semi-Lagrangian scheme of Algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le} presented in section \ref{sec:accuracy}, i.e. we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by equation \eqref{eq:velo_enright}.

To understand the effect of the CFL number on the scalability of the algorithm we perform one step of the semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$. We also perform the test for two different initial girds, a small grid with maximum level $l_\text{max} = 10$ and a large grid with maximum level $l_\text{max} = 12$. In both cases, the minimum level is $l_\text{min}=0$. After one advection step, these grids have approximately 15M and 255M nodes, respectively. 
\begin{table}[htbp]
	\begin{minipage}{.48\linewidth}
		\begin{center}
		\begin{tabular}{|c|cccccc|}
			\hline
				\diagbox{CFL}{\#p} & 16 & 32 & 64 & 128 & 256 & 512 \\
				\hline
				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
			\hline
		\end{tabular}
		\caption*{(a) $l_\text{max} = 10$}
		\end{center}
	\end{minipage}%
	\begin{minipage}{.48\linewidth}
		\begin{center}
		\begin{tabular}{|c|cccccc|}
			\hline
				\diagbox{CFL}{\#p} & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
			\hline
				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
			\hline
		\end{tabular}
		\caption*{(b) $l_\text{max} = 12$}
		\end{center}
	\end{minipage}	
	\caption{Number of sub-iterations required for the grid construction in Algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively. Note how the sub-iteration count increases with the CFL number but is almost independent of the number of processes. The slight dependence between the number of sub-iterations and the number of processes is most likely due to the dependence of round-off errors on the number of processes. Nonetheless, close examination of the Octrees generated (data not shown) reveals that they are identical and independent of the number of processes used to perform the test.}
	\label{tab:semilagrangian}
\end{table}
